---
title: "example display table"
author: "john"
format: 
  html:
    df-print: paged
editor: source
---

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(gt)
```

```{r}
economics |> 
  sample_n(25)
```

pivot longer so that I can gather summary statistics

```{r}
economics |>  
  mutate(my_era = lubridate::year(date) < 1980) |> 
  pivot_longer(cols = pce:unemploy, names_to = "my_variable", values_to = "value") 

```

gather summary statistics

```{r}
economics |>
  mutate(my_era = lubridate::year(date) < 1980) |> 
  pivot_longer(cols = pce:unemploy, names_to = "my_variable", values_to = "value") |> 
  group_by(my_era, my_variable) |> 
  summarise(my_mean = mean(value)) 
```

pivot wider so that a data frame can be printed in a specific manner

```{r}
economics |>
  mutate(my_era = lubridate::year(date) < 1980) |> 
  pivot_longer(cols = pce:unemploy, names_to = "my_variable", values_to = "value") |> 
  group_by(my_era, my_variable) |> 
  summarise(my_mean = mean(value)) |> 
  pivot_wider(names_from = my_era, values_from = my_mean)
```

print data frame in a specific manner

```{r}
economics |> 
  mutate(my_era = lubridate::year(date) < 1980) |> 
  pivot_longer(cols = pce:unemploy, names_to = "my_variable", values_to = "value") |> 
  group_by(my_era, my_variable) |> 
  summarise(my_mean = mean(value)) |> 
  pivot_wider(names_from = my_era, values_from = my_mean)
```


## editing tables with {gt}

this is how you make the table pretty.  If you need LaTeX or PDF documents, use similar concepts as {gt} but you'll youse {kableExtra} or {flextable} isntead.  See `basis_pivot_sumarize_print_kablePDF.qmd` for more information.

https://gt.rstudio.com/  is really well documented

```{r}
economics |> 
  mutate(my_era = if_else(lubridate::year(date) < 1980, "Xgen", "postgen")) |>   
  pivot_longer(cols = pce:unemploy, names_to = "my_variable", values_to = "value") |> 
  group_by(my_era, my_variable) |> 
  summarise(my_mean = mean(value), .groups = "drop") |> 
  # mutate(my_era = if_else(my_era == "postgen", 1, 0)) |> 
  pivot_wider(names_from = my_era, values_from = my_mean) |> 
  gt(rowname_col = "my_variable") |> 
  tab_header(
    title = "my special table",
    subtitle = "about economics data"
  ) |> 
  tab_stubhead(label = "Indicator") |> 
  tab_spanner(
    label = "X-era",
    columns = c(Xgen, postgen)
  )
```

