---
title: "Untitled"
format: 
  pdf:
    df-print: kable
editor: source
---

** Especially on Windows, run `quarto install tinytex` in the terminal.  This may take a while.  More info here: https://quarto.org/docs/output-formats/pdf-engine.html

MacOS and Unix may have different requirements.

ALSO:  Use {kableExtra} to stylize and manipulate your table (headers, subtitles, totals, LaTeX features, etc.)
Alternative to {kableExtra} is {flextable}  See more at https://davidgohel.github.io/flextable/reference/index.html    AND  
https://ardata-fr.github.io/flextable-book/

LAST bit of advice.  I'm not a regulare PDF document producer and hence I'm not familair with all that {kableExtra} or {flextable} can do.  But what they do is similar to {gt}.  The difference is that {gt} output is not suitable for PDF.  Inanycase, since i am familar with {gt} I will show you some conceptual table editing.  Youc an figure out how you do that with {kableExtra} or {flextable}.  I suspect {flextable} has better documentation.  See my example at the bottom of `basis_pivot_sumarize_print.qmd`.  In all cases, here and above, besure to render the document to get the full effect.

```{r}
library(tidyverse)
```

here's a summary

```{r}
starwars |> 
  select(name:hair_color) |> 
  slice_head(n = 8)
```


```{r}
my_special_df <- starwars |> 
  drop_na(mass, height) |> 
  mutate(species = fct_lump_min(species, 3)) |> 
  select(name:hair_color, species) |>  
  group_by(species) |> 
  summarise(mean(height), sd(mass))
my_special_df
```

```{r}
economics |> 
  sample_n(25)
```

pivot longer so that I can gather summary statistics

```{r}
economics |>  
  mutate(my_era = lubridate::year(date) < 1986) |> 
  pivot_longer(cols = pce:unemploy, names_to = "my_variable", values_to = "value") |> 
  sample_n(10)

```

gather summary statistics

```{r}
economics |>
  sample_n(25) |> 
  mutate(my_era = lubridate::year(date) < 1986) |> 
  pivot_longer(cols = pce:unemploy, names_to = "my_variable", values_to = "value") |> 
  group_by(my_era, my_variable) |> 
  summarise(my_mean = mean(value)) 
```

pivot wider so that a data frame can be printed in a specific manner

```{r}
economics |>
  sample_n(25) |> 
  mutate(my_era = lubridate::year(date) < 1986) |> 
  pivot_longer(cols = pce:unemploy, names_to = "my_variable", values_to = "value") |> 
  group_by(my_era, my_variable) |> 
  summarise(my_mean = mean(value)) |> 
  pivot_wider(names_from = my_era, values_from = my_mean)
```

print data frame in a specific manner

```{r}
economics |> 
  sample_n(25) |> 
  mutate(my_era = lubridate::year(date) < 1986) |> 
  pivot_longer(cols = pce:unemploy, names_to = "my_variable", values_to = "value") |> 
  group_by(my_era, my_variable) |> 
  summarise(my_mean = mean(value)) |> 
  pivot_wider(names_from = my_era, values_from = my_mean)
```
